The Multi-Stencil Language

### 2016-02-01 ARTICLES/IJPP/article/article.tex
article IJPP corrections Helene 29/01/2016
article IJPP corrections+example Helene 01/02/2016

\author{Helene Coullon         \and
        Chirtsian Perez \and
        Julien Bigot
}

@@

%------------------------------------------------------------------------------
\section{Introduction}
\label{sect:introduction}
Contributions :
\begin{itemize}
\item Formalism of a multi-stencil program and its parallelization
\item The Multi-stencil Language: agnostic from the type of mesh and of the choosen implementation
\item A first implementation of the MSL compiler (static scheduling + fusion + dump to SkelGIS/components)
\item Evaluation of the different types of parallelism introduced + compiler + fusion
\end{itemize}
%------------------------------------------------------------------------------
\section{Computational model of Multi-Stencil Programs}
\label{sect:formalism}
\input{formalism.tex}
% Formalism :
% \begin{itemize}
% \item mesh / mesh entities and groups / computations domains
% \item data and scalars
% \item computations : stencil, local and reduction
% \item Multi-Stencil Program
% \end{itemize}
%------------------------------------------------------------------------------
\section{Parallelization of Multi-Stencil Programs}
\label{sect:parallelism}
\input{parallelism.tex}
% Parallelization formalism :
% \begin{itemize}
% \item Data parallelism (introduction of synchronizations)
% \item Task parallelism (computation of the dependency graph)
% \end{itemize}
%------------------------------------------------------------------------------
\section{The Multi-Stencil Language}
\label{sect:msl}
\input{msl.tex}
The parallel empty skeleton of the application can be computed automatically from the language :
\begin{itemize}
\item grammar
\item example from the language to $\Gamma$
\end{itemize}
%------------------------------------------------------------------------------
\section{Static Scheduling}
\label{sect:msp}
\input{msp.tex}
% From the dependency graph we study a first dump: static scheduling
% \begin{itemize}
% \item description of the algorithm to get the binary tree decomposition of a minimal serie-parallel graph
% \item use of the serie-parallel tree decomposition to detect fusions of kernels
% \item example (the same one) from the dependency graph to the binary tree
% \end{itemize}
%------------------------------------------------------------------------------
\section{MSL Compiler}
\label{sect:comp}
\input{compiler.tex}
\begin{itemize}
\item Overall algrotihm~\ref{alg:compiler}
\item Big example from the language to the static sheduling/optimization~\ref{fig:example}
\end{itemize}
%------------------------------------------------------------------------------
\section{Evaluation}
\label{sect:eval}
\begin{itemize}
\item short description of the implementation (compiler in Python, DDS = SkelGIS, Components)
\item evaluation of the compiler (execution times of the different steps)
\item evaluation of the data parallelism (weak/strong scaling)

@@

\end{itemize}
%------------------------------------------------------------------------------
\section{Related work}
\label{sect:rel}
\begin{itemize}
\item Lizst, Pochoir, PATUS for optimizations/parallelization of a single stencil kernel + mesh dependent
\item Halide for pipeline of stencil computations = mesh dependent
\item Reuse of external solutions as SkelGIS, GA, Lizst, Pochoir etc.
\end{itemize}
%------------------------------------------------------------------------------
\section{Conclusion}
\label{sect:concl}
%------------------------------------------------------------------------------

>>>>

\author{Helene Coullon         \and
        Christian Perez \and
        Julien Bigot
}

@@

%------------------------------------------------------------------------------
\section{Introduction}
\label{sect:introduction}
Contributions :
\begin{itemize}
\item Formalism of a multi-stencil program and its parallelization
\item The Multi-stencil Language: agnostic from the type of mesh and of the choosen implementation
\item A first implementation of the MSL compiler (static scheduling + fusion + dump to SkelGIS/components)
\item Evaluation of the different types of parallelism introduced + compiler + fusion
\end{itemize}
%------------------------------------------------------------------------------
\section{Computational model of Multi-Stencil Programs}
\label{sect:formalism}
\input{formalism.tex}
% Formalism :
% \begin{itemize}
% \item mesh / mesh entities and groups / computations domains
% \item data and scalars
% \item computations : stencil, local and reduction
% \item Multi-Stencil Program
% \end{itemize}
%------------------------------------------------------------------------------
\section{Parallelization of Multi-Stencil Programs}
\label{sect:parallelism}
\input{parallelism.tex}
% Parallelization formalism :
% \begin{itemize}
% \item Data parallelism (introduction of synchronizations)
% \item Task parallelism (computation of the dependency graph)
% \end{itemize}
%------------------------------------------------------------------------------
\section{The Multi-Stencil Language}
\label{sect:msl}
\input{msl.tex}
The parallel empty skeleton of the application can be computed automatically from the language :
\begin{itemize}
\item grammar
\item example from the language to $\Gamma$
\end{itemize}
%------------------------------------------------------------------------------
\section{Static Scheduling}
\label{sect:msp}
\input{msp.tex}
% From the dependency graph we study a first dump: static scheduling
% \begin{itemize}
% \item description of the algorithm to get the binary tree decomposition of a minimal serie-parallel graph
% \item use of the serie-parallel tree decomposition to detect fusions of kernels
% \item example (the same one) from the dependency graph to the binary tree
% \end{itemize}
%------------------------------------------------------------------------------
\section{MSL Compiler}
\label{sect:comp}
\input{compiler.tex}
% \begin{itemize}
% \item Overall algrotihm~\ref{alg:compiler}
% \item Big example from the language to the static sheduling/optimization~\ref{fig:example}
% \end{itemize}
%------------------------------------------------------------------------------
\section{Evaluation}
\label{sect:eval}
\begin{itemize}
\item short description of the implementation (compiler in Python, DDS = SkelGIS, Components)
\item evaluation of the compiler (execution times of the different steps)
\item evaluation of the data parallelism (weak/strong scaling)

@@

\end{itemize}
%------------------------------------------------------------------------------
\section{Related work}
\label{sect:rel}
\begin{itemize}
\item Cartesian meshes : Pochoir, PATUS for optimizations/parallelization of a single stencil kernel
\item Unstructured meshes : Lizst for optimizations/parallelization of a single stencil kernel
\item Halide for pipeline of stencil computations in image (Cartesian) = mesh dependent
\item SYCL same Halide mais mesh dependent
\item Reuse of external solutions as SkelGIS, GA, Lizst, Pochoir etc.
\end{itemize}
%------------------------------------------------------------------------------
\section{Conclusion}
\label{sect:concl}
%------------------------------------------------------------------------------

### 2016-02-01 ARTICLES/IJPP/article/compiler.tex
article IJPP corrections+example Helene 01/02/2016

The overall compiler algorithm is represented in Algorithm~\ref{alg:compiler}. First, the parsing of the input file (written with the MSL language) gave enough information to build $\Gamma$, the ordered list of computations. Second, $\Gamma_{sync}$ and $\Gamma_{dep}$ can be built from $\Gamma$ using the recursive functions $F_{sync}$ and $F_{dep}$. Third, N-Shapes are removed from $\Gamma_{dep}$ to get a MSPD graph, and to be able to create its serie-parallel binary tree decomposition $TSP$ (line 6). Lines 7 to 9 are explained in the next section. Finally a dump of $TSP$ is performed by the compiler and will be described in Section~\ref{sect:eval}.

@@

\begin{myth}
\begin{itemize}
\item A sequence of sequence is a sequence if the order is kept. Thus for two sequence sub-trees $T_i$ and $T_j$, if $T_j \in V_i$, $T_i$ can be replaced by 
\begin{equation*}
T_i'(V_i \setminus root(T_j),E_i \cup replace(E_j,root(T_j),root(T_i)))
\end{equation*}
\item A parallel sub-tree of parallel sub-tree is a parallel sub-tree in any order. Thus for two parallel sub-trees $T_i$ and $T_j$, if $T_j \in V_i$, $T_i$ can be replaced by 
\begin{equation*}
T_i'(V_i \setminus root(T_j),E_i \cup replace(E_j,root(T_j),root(T_i)))
\end{equation*}
\end{itemize}
\end{myth}

@@

MSL code~\ref{fig:exmsl}. 
\begin{equation*}
\Gamma = [k_0,k_1,k_2,k_3,k_4,k_0,k_6,k_7,k_8]
\end{equation*}

\begin{equation*}
\Gamma_{sync} = [k_0,k_{0;1}^{sync},k_1,k_2,k_3,k_{1;4}^{sync},k_4,k_0,k_6,k_7,k_{7;8}^{sync},k_8]
\end{equation*}

where $k_{0;1}^{sync}$...

$\Gamma_{dep}$~\ref{} and $L^{-1}(\Gamma_{dep})$~\ref{}

% Using those additional reduction rules, the binary series-parallel tree decomposition can be computed in linear time. 
A complete example of reduction is given in Figure~\ref{fig:example}. The initial $\Gamma_{hybrid}$ DAG and its inverse line graph are given first, followed by the set of reductions. The label of the final single edge corresponds to the binary tree decomposition of the initial $\Gamma_{hybrid}$ graph.

%====BIG EXAMPLE
\begin{figure}[ht!]
\begin{center}
% INITIAL GRAPH
\subfloat[]{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[] (c0) at (0,0) {$k_0$};
   \node[] (star1) at (1,0) {$k_{0;1}^{sync}$};
   \node[] (c1) at (2,0) {$k_1$};
   \node[] (c2) at (3,0.5) {$k_2$};
   \node[] (star4) at (3,1.5) {$k_{1;4}^{sync}$};
   \node[] (c3) at (3,-0.5) {$k_3$};
   \node[] (c4) at (4,0.5) {$k_4$};
   \node[] (c5) at (4,-0.5) {$k_5$};
   \node[] (c6) at (5,0.5) {$k_6$};
   \node[] (c7) at (6,0) {$k_7$};
   \node[] (star8) at (7,0) {$k_{7;8}^{sync}$};
   \node[] (c8) at (8,0) {$k_8$};

  \path[->]
    (c0) edge node {} (star1)
    (star1) edge node {} (c1)
    (c1) edge node {} (c2)
         edge node {} (c3)
         edge node {} (star4)
    (star4) edge node {} (c4)
    (c2) edge node {} (c4)
    (c4) edge node {} (c6)
    (c3) edge node {} (c5)
    (c5) edge node {} (c7)
    (c6) edge node {} (c7)
    (c7) edge node {} (star8)
    (star8) edge node {} (c8);
  \end{tikzpicture}
}
\\
% INVERSE LINE GRAPH
\subfloat[]{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[circle,draw=black,fill=black,scale=0.3] (c0s) at (0,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c0d) at (1,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (star1d) at (2,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c1d) at (3,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c2d) at (4,0.5) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c3d) at (4,-0.5) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c4d) at (5,0.5) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c6d) at (6,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c7d) at (7,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (star8d) at (8,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c8d) at (9,0) {};

@@

\begin{figure}[ht!]
\begin{center}
% TRANSFORMATIONS
\subfloat[]{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[circle,draw=black,fill=black,scale=0.3] (c0s) at (0,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c1d) at (3,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c2d) at (4,0.5) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c3d) at (4,-0.5) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c4d) at (5,0.5) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c6d) at (6,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c8d) at (9,0) {};

@@

    %reduction c0 c1
    (s0) edge node {} (s1)
         edge node {} (c1)
    (s1) edge node {} (c0)
         edge node {} (star1)
    %reduction c7 c8
    (s2) edge node {} (s3)
         edge node {} (c8)
    (s3) edge node {} (c7)
        edge node {} (star8);
  \end{tikzpicture}
}
\\
\subfloat[]{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[circle,draw=black,fill=black,scale=0.3] (c0s) at (0,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c1d) at (3,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c2d) at (4,0.5) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c4d) at (5,0.5) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c6d) at (6,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c8d) at (9,0) {};

@@

\begin{figure}[ht!]
\begin{center}
\subfloat[]{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[circle,draw=black,fill=black,scale=0.3] (c0s) at (0,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c1d) at (3,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c6d) at (6,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c8d) at (9,0) {};

@@

    %reduction c2 c6
    (s6) edge node {} (p0)
         edge node {} (s7)
    %reduction c2 *4
    (p0) edge node {} (c2)
         edge node {} (star4)
    (s7) edge node {} (c4)
         edge node {} (c6);
  \end{tikzpicture}
}
\\
\subfloat[]{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[circle,draw=black,fill=black,scale=0.3] (c0s) at (0,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c1d) at (3,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c6d) at (7,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c8d) at (9,0) {};

@@

    (p1) edge node {} (s6)
         edge node {} (s5)
    %reduction c3 c5
    (s5) edge node {} (c3)
         edge node {} (c5)
    %reduction c2 c6
    (s6) edge node {} (p0)
         edge node {} (s7)
    %reduction c2 *4
    (p0) edge node {} (c2)
         edge node {} (star4)
    (s7) edge node {} (c4)
         edge node {} (c6);
  \end{tikzpicture}
}
\\
\subfloat[]{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[circle,draw=black,fill=black,scale=0.3] (c0s) at (0,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c8d) at (9,0) {};

>>>>

The overall compiler algorithm is represented in Algorithm~\ref{alg:compiler}. First, the parsing of the input file (written with the MSL language) gave enough information to build $\Gamma$, the ordered list of computation kernels. Second, $\Gamma_{sync}$ and $\Gamma_{dep}$ can be built from $\Gamma$ using the recursive functions $F_{sync}$ and $F_{dep}$. Third, N-Shapes are removed from $\Gamma_{dep}$ to get a MSPD graph, and to be able to create its serie-parallel binary tree decomposition $TSP$ (line 6). Lines 7 to 9 are explained in the next section. Finally a dump of $TSP$ is performed by the compiler and will be described in Section~\ref{sect:eval}.

@@

\begin{myth}
A sequence of sequence is a sequence if the order is kept. Thus for two sequence sub-trees $T_i$ and $T_j$, if $T_j \in successors(root(T_i))$, $T_i$ can be replaced by 
\begin{equation*}
T_i'(V_i \setminus root(T_j),E_i \cup replace(E_j,root(T_j),root(T_i)))
\end{equation*}
\end{myth}

\begin{myth}
A parallel sub-tree of parallel sub-tree is a parallel sub-tree in any order. Thus for two parallel sub-trees $T_i$ and $T_j$, if $T_j \in successors(root(T_i))$, $T_i$ can be replaced by 
\begin{equation*}
T_i'(V_i \setminus root(T_j),E_i \cup replace(E_j,root(T_j),root(T_i)))
\end{equation*}
\end{myth}

@@

Figure~\ref{fig:exmsl} gives an example of $\mathcal{MSP}$ program. From this example, and using a very simple parser, the following ordered list of computation kernels can be extracted, including all information on mesh, entities, computation domains and data:
\begin{equation*}
\Gamma = [k_0,k_1,k_2,k_3,k_4,k_0,k_6,k_7,k_8]
\end{equation*}

From this ordered list of computation kernels $\Gamma$ and included information, synchronizations can be automatically detected from the call to $F_{sync}(\Gamma,0)$ to get the synchronized ordered list of kernels:

\begin{equation*}
\Gamma_{sync} = [k_0,k_{0;1}^{sync},k_1,k_2,k_3,k_{1;4}^{sync},k_4,k_0,k_6,k_7,k_{7;8}^{sync},k_8],
\end{equation*}

\begin{subequations}
where
\begin{align}
        k_{0;1}^{sync}=(\emptyset,\{(B,\bigcup_{e \in D_0} nce(e))\},(B,\bigcup_{e \in D_0} nce(e)),identity),\\
        k_{1;4}^{sync}=(\emptyset,\{(C,\bigcup_{e \in D_1} nec(e))\},(C,\bigcup_{e \in D_1} nec(e)),identity),\\
        k_{7;8}^{sync}=(\emptyset,\{(I,\bigcup_{e \in D_7} ncc(e))\},(I,\bigcup_{e \in D_7} ncc(e)),identity).
\end{align}
\end{subequations}

From $\Gamma_{sync}$, the dependency DAG can be built. For example, as $k_0$ computes $B$ and $k_1$ reads $B$, $k_0$ and $k_1$ becomes vertices of $\Gamma_{dep}$, and an arc $(\overset{\frown}{k_0,k_1})$ is added to $\Gamma_{dep}$. The overall $\Gamma_{dep}$ built from the call to $F_{dep}(\Gamma_{sync},0)$ is drawn in Figure~\ref{fig:depdep}. $L^{-1}(\Gamma_{dep})$ also represented in Figure~\ref{fig:depline} is the line digraph of $\Gamma_{dep}$.

%====BIG EXAMPLE
\begin{figure}[ht!]
\begin{center}
% INITIAL GRAPH
\subfloat[$\Gamma_{dep}$\label{fig:depdep}]{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[] (c0) at (0,0) {$k_0$};
   \node[] (star1) at (1,0) {$k_{0;1}^{sync}$};
   \node[] (c1) at (2,0) {$k_1$};
   \node[] (c2) at (3,0.5) {$k_2$};
   \node[] (star4) at (3,1.5) {$k_{1;4}^{sync}$};
   \node[] (c3) at (3,-0.5) {$k_3$};
   \node[] (c4) at (4,0.5) {$k_4$};
   \node[] (c5) at (4,-0.5) {$k_5$};
   \node[] (c6) at (5,0.5) {$k_6$};
   \node[] (c7) at (6,0) {$k_7$};
   \node[] (star8) at (7,0) {$k_{7;8}^{sync}$};
   \node[] (c8) at (8,0) {$k_8$};

  \path[->]
    (c0) edge node {} (star1)
    (star1) edge node {} (c1)
    (c1) edge node {} (c2)
         edge node {} (c3)
         edge node {} (star4)
    (star4) edge node {} (c4)
    (c2) edge node {} (c4)
    (c4) edge node {} (c6)
    (c3) edge node {} (c5)
    (c5) edge node {} (c7)
    (c6) edge node {} (c7)
    (c7) edge node {} (star8)
    (star8) edge node {} (c8);
  \end{tikzpicture}
}
\\
% INVERSE LINE GRAPH
\subfloat[$L^{-1}(\Gamma_{dep})$\label{fig:depline}]{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[circle,draw=black,fill=black,scale=0.3] (c0s) at (0,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c0d) at (1,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (star1d) at (2,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c1d) at (3,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c2d) at (4,0.5) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c3d) at (4,-0.5) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c4d) at (5,0.5) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c6d) at (6,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c7d) at (7,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (star8d) at (8,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c8d) at (9,0) {};

@@

Figure~\ref{fig:SPS} illustrates two binary tree decomposition following the sequence rule of Figure~\ref{fig:rules}. This rule is first applied to $k_0$ and $k_{0;1}^{sync}$, then to the resulting tree to $k_1$. The same is applied onto $k_7$ and $k_{7;8}^{sync}$ and then to the resulting tree to $k_8$. In Figure~\ref{fig:SPP} the parallel rule of Figure~\ref{fig:rules} is applied onto $k_{1;4}^{sync}$ and $k_2$.

\begin{figure}[ht!]
\begin{center}
% TRANSFORMATIONS
\subfloat[Sequence rule applied onto $k_0$ $k_{0;1}^{sync}$ and $k_1$, and onto $k_7$ $k_{7;8}^{sync}$ and $k_8$\label{fig:SPS}]{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[circle,draw=black,fill=black,scale=0.3] (c0s) at (0,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c1d) at (3,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c2d) at (4,0.5) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c3d) at (4,-0.5) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c4d) at (5,0.5) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c6d) at (6,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c8d) at (9,0) {};

@@

    %reduction c0 c1
    (s0) edge node {} (s1)
         edge node {} (c1)
    (s1) edge node {} (c0)
         edge node {} (star1)
    %reduction c7 c8
    (s2) edge node {} (s3)
         edge node {} (c8)
    (s3) edge node {} (c7)
        edge node {} (star8);
  \end{tikzpicture}
}
\\
\subfloat[Parallel rule applied onto $k_{1;4}^{sync}$ and $k_2$\label{fig:SPP}]{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[circle,draw=black,fill=black,scale=0.3] (c0s) at (0,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c1d) at (3,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c2d) at (4,0.5) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c4d) at (5,0.5) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c6d) at (6,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c8d) at (9,0) {};

@@

Finally, the rest of the successive transformations, following rules of Figure~\ref{fig:rules}, are illustrated in Figure~\ref{fig:example} until a single edge is reach which ends the algorithm.

\begin{figure}[ht!]
\begin{center}
\subfloat{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[circle,draw=black,fill=black,scale=0.3] (c0s) at (0,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c1d) at (3,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c6d) at (6,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c8d) at (9,0) {};

@@

    %reduction c2 c6
    (s6) edge node {} (p0)
         edge node {} (s7)
    %reduction c2 *4
    (p0) edge node {} (c2)
         edge node {} (star4)
    (s7) edge node {} (c4)
         edge node {} (c6);
  \end{tikzpicture}
}
\\
\subfloat{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[circle,draw=black,fill=black,scale=0.3] (c0s) at (0,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c1d) at (3,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c6d) at (7,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c8d) at (9,0) {};

@@

    (p1) edge node {} (s6)
         edge node {} (s5)
    %reduction c3 c5
    (s5) edge node {} (c3)
         edge node {} (c5)
    %reduction c2 c6
    (s6) edge node {} (p0)
         edge node {} (s7)
    %reduction c2 *4
    (p0) edge node {} (c2)
         edge node {} (star4)
    (s7) edge node {} (c4)
         edge node {} (c6);
  \end{tikzpicture}
}
\\
\subfloat{
\begin{tikzpicture}[shorten >=1pt, node distance=2cm, on grid, auto]
   \node[circle,draw=black,fill=black,scale=0.3] (c0s) at (0,0) {};
   \node[circle,draw=black,fill=black,scale=0.3] (c8d) at (9,0) {};
